---
title: "Class14"
author: "Khoi Tran"
date: "May 16, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE, row.names = 1)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

# of genes
```{r}
nrow(counts)
```

Check if correspondence of out colData col and our meta data rows are the same
```{r}
colnames(counts)
metadata$id
```
```{r}
(colnames(counts) == metadata$id)
```

```{r}
head(counts)
head(metadata)
```

```{r}
control.inds <- metadata[,"dex"]=="control"
control <- metadata[metadata[,"dex"]=="control",]
control

```
Lining up Meta Data and Scaled Data
```{r}
head(counts[,control$id])
control.mean <- rowSums(counts[,control$id]) / length(control$id)
```

```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated.mean <- rowSums(counts[,treated$id]) / length(treated$id)
```

```{r}
meancounts <- data.frame(control.mean, treated.mean)
colSums(meancounts)
plot(meancounts)
```
 
Log Scale of Graph
```{r}
plot(meancounts, log="xy")
```

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```
Removing 0 values
```{r}
to.rm <- unique(which(meancounts[,1:2] == 0, arr.ind = TRUE)[,1])
```

```{r}
mycounts <- meancounts[-to.rm,]
nrow(mycounts)
```

How many genes are up regulated
```{r}
up.inds <- mycounts$log2fc > 2
sum(up.inds)
```

How many genes are down regulated
```{r}
down.inds <- mycounts$log2fc < -2
sum(down.inds)
```

Finding Gene Name and Function (Annotating Data)
```{r}
anno <- read.csv("annotables_grch38.csv")
head(mycounts[up.inds,])
head(anno)
```

Using Merge to add the annotation data to RNA seq data
```{r}
merge(mycounts,anno, by.x = "row.names", by.y = "ensgene")
```

Getting Databases
```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```
```{r}
columns(org.Hs.eg.db)
```

Using mapIds:
ENSEMBL is the format of gene names
```{r}
mycounts$symbols <- mapIds(org.Hs.eg.db, keys = row.names(mycounts), keytype = "ENSEMBL", column = "SYMBOL")
head(mycounts)
```

```{r}
mycounts$entrez <- mapIds(org.Hs.eg.db, keys = row.names(mycounts), keytype = "ENSEMBL", column = "ENTREZID")
head(mycounts)
```

Using DeSeq2
```{r}
library(DESeq2)
counts2 <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
dds <- DESeqDataSetFromMatrix(countData=counts2, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
res <- results(dds)
summary(res)
res05 <- results(dds, alpha = 0.05)
summary(res05)
```

```{r}
res01 <- as.data.frame(results(dds, alpha = 0.01))
head(res01)
```

Ordering Results
```{r}
ord.inds <- order( res01$padj )
head(res01[ord.inds,])
write.csv( res01[ord.inds,], file = "signif01_results.csv")
```

Volcano Plot
```{r}
plot(res01$log2FoldChange, res01$padj)
plot(res01$log2FoldChange, -log(res01$padj))
```

```{r}
mycols <- rep("orange", nrow(res01))
mycols[res01$padj < 0.01] <- "red"
mycols[ abs(res01$log2FoldChange) > 2] <- "blue"
mycols[(res01$padj< 0.01) & abs(res01$log2FoldChange) > 2] <- "green"
plot(res01$log2FoldChange, -log(res01$padj), 
     xlab = "Drug/Control (-log)",
     ylab = "P-value (-log)",
     col = mycols)
```





